---
title: "HW2 EDA"
author: "Zhanglin Shangguan"
date: "2022/2/10"
output:
  pdf_document: default
  html_document:
    df_print: paged
---
### Data Cleaning

```{r}
setwd('C:/Users/SGZL/Desktop/Exploratory Analytics/SunCountry_data')
library(dplyr)
library(tidyr)
library(cluster)
library(clustMixType)
library(ggplot2)
library(lubridate)

df <- read.csv('SunCountry.csv')

## Filtering out extreme age groups, ungendered groups
## The Airlines with which booking was made is SY
## Combining UflyMemberStatus and CardHolder together to create new column Status
df <- df %>%
  filter(Age > 0 & Age < 110 & MarketingAirlineCode == "SY" & GenderCode != "U") %>% 
  unite("Status", UflyMemberStatus:CardHolder, remove = TRUE)

## Categorize Status into 5 groups
df$Status[which(df$Status == "_")] = "Non-Member"
df$Status[which(df$Status == "Standard_false")] = "Standard"
df$Status[which(df$Status == "Standard_true")] = "Standard Carholder"
df$Status[which(df$Status == "Elite_false")] = "Elite"
df$Status[which(df$Status == "Elite_true")] = "Elite Cardholder"


df <- df %>%
  mutate(EnrollDate = ymd_hms(EnrollDate), ServiceStartDate = ymd(ServiceStartDate), 
         PNRCreateDate = ymd(PNRCreateDate)) %>% # Convert dates into date format
  mutate(EnrollDate = floor_date(EnrollDate, "day")) %>% #Convert datetime into date format
  # The number of days of memebership before the flight
  mutate(memebership_duration_before_flight = difftime(ServiceStartDate, EnrollDate, units = "days"),
         # How many days in advance that customer booked their trip before traveling
         Daysinadvance_booked_flight = difftime(ServiceStartDate, PNRCreateDate, units = "days")) %>%
  #The month and the day of the week of flight, 1 is Sunday for day_of_week_flight
  mutate(month_flight = month(ServiceStartDate), day_of_week_flight = wday(ServiceStartDate, abbr = FALSE)) %>%
  # If customer have discounts, 0 - no discount, 1 - disount
  mutate(discount = if_else(BookedProduct == "", 0, 1)) %>%
  # Creating Stoppover column to indicate the duration
  mutate(Stopover = case_when(
    StopoverCode == "" ~ "Direct Flight",
    StopoverCode == "O" ~ "Layovers",
    StopoverCode == "X" ~ "24hrs Stopover")) %>%
  # unique Id
  unite("Unique_ID", c(EncryptedName, birthdateid, TicketNum) ,remove = FALSE)


```

```{r}
df %>%
  filter(Unique_ID == "41414B48555344696420493F7C206765742074686973207269676874414C4C49534F4E20414E4E_55353_3372107204347")
```


```{r}
#Frequqent Flyer

df1 = df %>% 
  group_by(Unique_ID) %>% 
  dplyr::summarise(countv = n())

#df2 = df1%>% order_by(countv, desc)
#

library(plyr)
df <- join(df, df1,type = "left", by ="Unique_ID")

## define frequent flyer
df <- df%>%  mutate(
      FrequentFlyer =
      case_when(
        countv >= 3 ~ "Frequent Flyer",
        TRUE ~ " Not Frequent Flyer")
      )

```

#### Converting variables into factors
```{r}
df$GenderCode = factor(df$GenderCode)
df$TrvldClassOfService = factor(df$TrvldClassOfService)
df$BookingChannel = factor(df$BookingChannel)
df$BkdClassOfService = factor(df$BkdClassOfService)
df$Stopover = factor(df$Stopover)
df$Status = factor(df$Status)
df$month_flight = factor(df$month_flight)
df$FrequentFlyer = factor(df$FrequentFlyer)
```


### EDA

```{r}
df %>%
  group_by(TrvldClassOfService) %>%
  dplyr::summarise(Avg_age = mean(Age)) %>%
  ggplot(aes(x = TrvldClassOfService, y = Avg_age)) + 
  geom_bar(stat = 'identity') + ylab("Average Age")
```
### T-test to test there is significant age difference between one who travel first class and coach. The results has shown that there is a significant difference between the age for Coach Class and First Class travelers.

```{r}
df_t <- df
df_t$TrvldClassOfService[which(df_t$TrvldClassOfService == "Discount First Class")] = "First Class"

t.test(Age ~ TrvldClassOfService, data = df_t)

```

### Check the distribution of top 4 most popular booking channel. 

```{r}
df %>%
  dplyr::count(BookingChannel) %>%
  arrange(desc(n)) %>%
  head(4) %>%
  ggplot(aes(x = BookingChannel, y = n)) + geom_bar(stat = 'identity') +
  xlab('Booking Chaneel')
  
  
```

### Fist class break down to study the demographic who often purchase first class
```{r}

df_ages = df %>% mutate(Age_disc = ifelse(Age <= 30, "Young", 
                           ifelse(Age > 30 & Age <=50,"Middle Age", "Old")))

ggplot(df_ages %>% 
         filter(TrvldClassOfService != 'Coach') %>% 
         group_by(Age_disc, TrvldClassOfService, GenderCode) %>% 
         dplyr::summarise(count = n()), aes(x=TrvldClassOfService, y=count, 
                                     groups=Age_disc, fill=Age_disc)) +                          geom_bar(stat='identity', position='dodge')  + 
         labs(title="First Class Breakdown by Gender and Age", 
              x="Class", y="Number of Flights", fill='Age Groups') +  
         facet_wrap(GenderCode~.)
```


```{r}
df_freq1 = df %>% group_by(discount, BkdClassOfService) %>% dplyr::summarise(discount1 = n())
df_freq1
p1 <- ggplot(df_freq1%>%filter(BkdClassOfService == "Coach"), aes(fill=factor(discount), x=BkdClassOfService, y = discount1)) +
    geom_bar(position="dodge", stat="identity")
p2 <- ggplot(df_freq1%>%filter(BkdClassOfService == "Discount First Class"), aes(fill=factor(discount), x=BkdClassOfService, y = discount1)) +
    geom_bar(position="dodge", stat="identity")
p3 <- ggplot(df_freq1%>%filter(BkdClassOfService == "First Class"), aes(fill=factor(discount), x=BkdClassOfService, y = discount1)) +
    geom_bar(position="dodge", stat="identity")
#df_freq[df_freq$Status == "Elite"]
```



```{r}
## We had to change plotly as it does not support pdf
##library(plotly)
##subplot(ggplotly(p1), ggplotly(p2), ggplotly(p3),nrows = 1)
require(gridExtra)
par(mfrow=c(1,3))
p1
p2
p3
```



### Kprototype Clustering for segementing customers
```{r}
library(clustMixType)
```


```{r}
normalize <- function(x){
  return ((x - min(x))/(max(x) - min(x)))}

set.seed(1000)

df_sample <- df %>%
  mutate(Daysinadvance_booked_flight = as.numeric(Daysinadvance_booked_flight)) %>%
  select("GenderCode", "Age", "TrvldClassOfService",
         "BaseFareAmt", "Status", "FrequentFlyer", "month_flight") %>%
  filter(BaseFareAmt != 0) %>%
  sample_n(30000, replace = FALSE) %>%
  mutate(BaseFareAmt_scaled = normalize(BaseFareAmt)) %>%
  mutate(Age_scaled = normalize(Age))
```


### Selecting clsutering features and perfom random sampling of 40000 records.
```{r}
normalize <- function(x){
  return ((x - min(x))/(max(x) - min(x)))}

set.seed(1000)

df_sample1 <- df %>%
  mutate(Daysinadvance_booked_flight = as.numeric(Daysinadvance_booked_flight)) %>%
  mutate(BaseFareAmt_scaled = normalize(BaseFareAmt)) %>%
  mutate(Age_scaled = normalize(Age)) %>%
  filter(BaseFareAmt != 0) %>%
  sample_n(40000, replace = FALSE)

df_sample1_use <- df_sample1 %>%
  select("GenderCode", "Age_scaled", "TrvldClassOfService",
         "BaseFareAmt_scaled", "Status", "FrequentFlyer", "month_flight")
```

### Descriptive statistics for the sample
```{r}
summary(df_sample1_use)
```


### Picking the number of clusters based on SSE, we pick 5.
```{r}
set.seed(9845)
SSE_curve <- c()
for (k in 1:8){
  kpro <- kproto(df_sample1_use, k)
  sse <- sum(kpro$withinss)
  SSE_curve[k] <- sse
}

plot(1:8, SSE_curve, type="b", xlab="Number of Clusters", ylab="SSE")

```

### Running clustering model
```{r}
set.seed(1000)
sy_kproto <- df_sample1_use %>%
  clustMixType::kproto(k = 4, nstart = 30)
```


```{r}
df_sample1_use_temp <- df_sample1 %>%
  select("GenderCode", "Age", "TrvldClassOfService",
         "BaseFareAmt", "Status", "FrequentFlyer", "month_flight")


df_sample1_use_temp <- 
  df_sample1_use_temp %>%
  mutate(kproto_cluster = sy_kproto$cluster)
table(df_sample1_use_temp$kproto_cluster)
```

Clustering plots
```{r}
clprofiles(sy_kproto, df_sample1_use_temp)
```

```{r}
summary(sy_kproto)
```

## Popular Destination
### For popular destination, we take out MSP from the visulization beucase it is
### obvious that MSP has the largest amount of traffic.

Cluster 1
```{r}
C1 <- df %>%
  filter(GenderCode == "M" & Status != "Non-Member") %>%
  filter(Age > 30 & Age < 50) %>%
  filter(month_flight == 3 | month_flight == 2) %>% 
  group_by(ServiceEndCity) %>% 
  dplyr::summarise(count = n(), .groups = 'drop') %>%
  arrange(desc(count)) %>%
  head(7)

```

### Top destinations for this cluster
```{r}
ggplot(C1 %>% 
         filter(ServiceEndCity != 'MSP'), 
         aes(x = reorder(ServiceEndCity, -count), y = count)) + 
         geom_bar(stat = "summary", position="dodge") + 
         labs(title="Top 6 Destinations for Cluster 1", 
         x=" Destination Airport", y="Count of flights") 
```


Cluster 2
```{r}
df$discount = factor(df$discount)

C2 <- df %>%
  filter(GenderCode == "F" & Status == "Non-Member" & Age < 30) %>%
  filter( month_flight == 7) %>% 
  group_by(ServiceEndCity) %>% 
  dplyr::summarise(count = n()) %>% 
  arrange(desc(count)) %>%
  head(7)
C2
```
### Top destinations for this cluster
```{r}

ggplot(C2 %>% 
         filter(ServiceEndCity != 'MSP'), 
         aes(x = reorder(ServiceEndCity, -count), y = count)) + 
         geom_bar(stat = "summary", position="dodge") + 
         labs(title="Top 6 Destinations for Cluster 2", 
              x=" Destination Airport", y="Count of flights") +  
         scale_x_discrete(guide = guide_axis(n.dodge=1)) +
         ylim(c(0,6000))
```



Cluster 3
```{r}

C3 <- df %>%
  filter(GenderCode == "F" & Status != "Non-Member") %>%
  filter(Age > 30 & Age < 50) %>%
  filter(month_flight == 12) %>% 
  group_by(ServiceEndCity) %>% 
  dplyr::summarise(count = n(), .groups = 'drop') %>%
  arrange(desc(count)) %>%
  head(7)

```


### Top destinations for this cluster
```{r}
ggplot(C3 %>% 
         filter(ServiceEndCity != 'MSP'),
         aes(x = reorder(ServiceEndCity, -count), y = count)) + 
         geom_bar(stat = "summary", position="dodge") + 
         labs(title="Top 6 Destinations for Cluster 3", 
         x=" Destination Airport", y="Count of flights") +  
         ylim(c(0,600))

```

Cluster 4
```{r}
C4 <- df %>%
  filter(GenderCode == "M" & Status != "Non-Member") %>%
  filter(Age < 30) %>%
  filter(month_flight == 8) %>% 
  group_by(ServiceEndCity) %>% 
  dplyr::summarise(count = n(), .groups = 'drop') %>%
  arrange(desc(count)) %>%
  head(7)
```

### Top destinations for this cluster
```{r}
ggplot(C4 %>% 
         filter(ServiceEndCity != 'MSP'), 
         aes(x = reorder(ServiceEndCity, -count), y = count)) + 
         geom_bar(stat = "summary", position="dodge") + 
         labs(title="Top 10 Destinations for Cluster 4", 
         x=" Destination Airport", y="Count of flights") +  
         ylim(c(0,500))
```


## Website Analysis 
```{r}
df=df%>%unite("Unique_ID", EncryptedName:birthdateid, remove = FALSE)
df$GenderCode=as.factor(df$GenderCode)
df$BookingChannel=as.factor(df$BookingChannel)
df$Status=as.factor(df$Status)
df$TrvldClassOfService=as.factor(df$TrvldClassOfService)
df$Stopover=as.factor(df$Stopover)
df$Daysinadvance_booked_flight=as.numeric(df$Daysinadvance_booked_flight)
df$memebership_duration_before_flight=as.numeric(df$memebership_duration_before_flight)
options(scipen = 100)


```


```{r}
set.seed(5)
df_web<-df%>%filter(BookingChannel=="SCA Website Booking")
df_web<-df_web%>%sample_n(15000,replace=FALSE)
df_cluster=df_web%>%select(Unique_ID,Age,TrvldClassOfService,Daysinadvance_booked_flight)
glimpse(df_cluster)
```

```{r}
library(ISLR) # for college dataset
library(cluster) # for similarity and pam
library(Rtsne) # for t-SNE plot
library(ggplot2) # for visualization
library(clustMixType) #for kprototypes
## calculating gower distance
gower_dist <- daisy(df_cluster[, -1],
                    metric = "gower",
                    type = list(logratio = 3))
summary(gower_dist)
gower_mat <- as.matrix(gower_dist)
```

```{r}
# to determine proper k value
sil_curve <- c()
for (k in 2:10) {
  pam_fit <- pam(gower_dist, diss = TRUE,k = k)
  #PAM interally computes the silhouette measure
  sil_curve[k] <- pam_fit$silinfo$avg.width
}
sil_curve = sil_curve[2:10]
plot(2:10, sil_curve, type="b", xlab="Number of Clusters", ylab="Silhouette")
num_pam_clusters = which.max(sil_curve)+1
```

### K-mediods clustering
```{r pam-sum}
pam_fit <- pam(gower_dist, diss = TRUE, k =5)

pam_results <- df_cluster %>%
  mutate(cluster = pam_fit$clustering) %>%
  group_by(cluster) %>%
  do(the_summary = summary(.))

### clustering result summary
pam_results$the_summary
```

### clustering results 
```{r}
pam_results
```

### #the size of each cluster
```{r}

dfM=df_cluster[pam_fit$medoids, ]
dfM
```
